Evaluative adjective sentences: \\ A question-based analysis of projection

### 2019-06-18 paper/previous-submissions/first-Glossa-submission.tex
added round 2 reviews, started new version of paper

EASs have traditionally been analyzed as presupposing the prejacent (e.g., \citealt{norrick78,barker02,oshima09b,kertz2010}). On such analyses,  evaluative adjectives like {\em stupid} lexically specify that the prejacent
must be entailed by or satisfied in the common ground of the interlocutors in
order for an EAS to be interpretable (e.g.,
\citealt{heim83,vds92}). Such analyses are motivated by and straightforwardly account for examples like those in (\ref{f2}a-d) in which the prejacent projects over entailment-canceling operators, i.e., negation in (\ref{f2}a), a polar question in (\ref{f2}b), the possibility adverb {\em perhaps} in (\ref{f2}c) and the antecedent of the conditional in (\ref{f2}d). That is, speakers who utter (\ref{f2}a-d) may be taken to be committed to the prejacent, that Feynman danced on the table, even though the evaluative adjective occurs under an entailment-canceling operator. The evaluation, on the other hand, that Feyman dancing on the table was stupid, is not projective content but denied in (\ref{f2}a), questioned in (\ref{f2}b), and merely considered a possibility in (\ref{f2}c) and (\ref{f2}d).

@@

We discuss below why such examples are problematic for analyses according to which the prejacent is a presupposition. But we first need to mention that even though all native speakers of American English appear to be able to retrieve the intended interpretations of examples like (\ref{nat}), a sizeable portion strongly prefers to realize such interpretations with variants that includes {\em enough}, as in {\em I wasn't stupid enough to go stumbling through the junkyard in the dark} for (\ref{nat}b); see also \citealt{karttunen2013} and \citealt{karttunen-etal2014} for this observation. Our goal in this paper is to analyze the interpretation of EASs and we assume that all native speakers of American English have access to both interpretations, that is, an interpretation according to which the prejacent projects, as in (\ref{f2}), and an interpretation according to which it doesn't project, as in (\ref{nat}). Evidence that examples in which the prejacent does not project are part of the productive language use of some native speakers of American English comes from the existence of naturally occurring examples like (\ref{nat}); further  such examples are presented in this paper. Additional evidence comes from the fact that the audiences to which this research was presented over the years always included native speakers that judged examples like (\ref{nat}) to be perfectly acceptable. More systematic evidence for the existence of such speakers comes from the acceptability rating study presented in Appendix \ref{s-acc}: of the 94 self-reported native speakers of American English that participated in the study, about 20-30\% judged negated evaluative adjective sentences without {\em enough} to be acceptable under a non-projecting interpretation of the prejacent. We return to the observed variation among native speakers of American English in section \ref{s5}.

@@

This behavior of the evaluation is not predicted by the two analyses that consider content other than the prejacent: \citealt{oshima09b} and \citealt{barker02}. In \citealt{oshima09b}, the prejacent is a presupposition and what is asserted is the following complex content: from the prejacent it can be inferred that the denotation of the subject is in the extension of the evaluative adjective (p.371). That is, (\ref{f2}a) is correctly predicted to presuppose that Feynman danced on the table and to assert that it cannot be inferred from Feynman dancing on the table that Feyman is stupid. Similarly then, (\ref{nat}b) is predicted to convey that the speaker did not stumble through the junkyard (the prejacent is locally accommodated) and to assert that it cannot be inferred from the speaker stumbling through the junkyard that the speaker is stupid. This assertion is too weak: crucially, it does not follow the speaker stumbling through the junkyard would be stupid.

@@

\citepos{karttunen-etal2014} analysis improves over previous analyses because it is able to derive both attested interpretations. But the analysis comes at a cost: first, evaluative adjectives are systematically ambiguous, and, second, as we argue in section \ref{s3}, it misses out on an empirical generalization about how common ground information  influences the interpretation of EASs. In this paper, we develop an analysis of EASs that accounts for the two interpretations with a single lexical entry for evaluative adjectives. The crux of the analysis is that evaluative adjectives do not lexically specify that the prejacent is a presupposition. In section \ref{s2}, we show, based on a study of naturally occurring NEASs, that the prejacent is not highly projective. We take this to provide empirical support for not analyzing the prejacent as a lexically-specified presupposition. In section \ref{s3}, we present an analysis of EASs according to which the projectivity of the prejacent of an utterance in which the evaluative adjective is embedded under an entailment-canceling operator is predicted by the extent to which the prejacent is at-issue with respect to the Discourse Question addressed by the utterance. We test three predictions of the analysis in section \ref{s4} and discuss two open issues in section \ref{s5}. The paper concludes in section \ref{s6}.

@@

\citet{karttunen-etal2014} provided naturally occurring examples that showed that the prejacent of NEASs need not project. The goal of the study we present in this section was to more systematically explore how projective the prejacent is in naturally occurring NEAS. Following \citealt{tbd-variability}, we assume that projectivity is a gradient property of utterance content: the more the speaker or author of the utterance is taken to be committed to the truth of an utterance content, the more projective the content is. To explore the projectivity of the prejacent, native speakers of American English were presented with naturally occurring NEAS and asked to rate the projectivity of the prejacent.\footnote{\label{f-git}The
materials, data and the R code for generating the figures and analyses of the experiments reported on in this paper are available at
\url{https://github.com/judith-tonhauser/evaluative-adjectives}.}

@@

Each target stimulus was attributed to an author who was identified by name, as shown for the sample stimuli in (\ref{target1}). As shown in the `Question to participants' presented with each stimulus in (\ref{target1}), we used the `certain that' diagnostic for projectivity to assess whether participants took the author to be committed to the prejacent, i.e., whether the prejacent projects over negation. For other applications of the `certain that' diagnostic for projection see \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017} and \citealt{tbd-variability}. In these `Questions to participants', the prejacent was realized as the finite complement clause of {\em certain}. The tense of the finite clause that realized the prejacent was determined by the temporal and aspectual properties of the
NEAS: it was past tense for past tense NEAS, as in (\ref{target1}a), and future tense for future tense NEAS, as in (\ref{target1}b); for non-past tense NEAS, the finite clause that realized the prejacent was realized in the non-past tense if the eventuality denoted by the prejacent was
stative, as in (\ref{target1}c), and a disjunction of a past tense and a future tense verb if the prejacent was eventive, as in (\ref{target1}d).

@@

In sum, our study of naturally occurring NEASs shows that the prejacent is not highly projective. These findings converge with those of \citet{tbd-variability}, who studied the projectivity of a variety of utterance contents and found that the prejacent of four (constructed) polar interrogative EASs with {\em stupid} was significantly less projective than the content of the clausal complement of the emotive predicate {\em be annoyed} and the cognitive predicate {\em know}. In the next section, we develop an analysis according to which the projectivity of the prejacent is not due to its lexical status as a presupposition but instead derived from the question that the utterance of the EAS is taken to address.

@@

@@

@@

The examples in (\ref{ai}) illustrate four Discourse Questions that utterances of EASs can address. These Discourse Questions show that both lexical entailments of EASs, the prejacent and the generalization, can be at-issue. The Discourse Questions in (\ref{ai}a) and (\ref{ai}b) are about the prejacent: both the question of whether Sam got a ticket in (\ref{ai}a) and who got a ticket in (\ref{ai}b) are answered by the prejacent of B's utterances, that Sam got a ticket.\footnote{Some native speakers of American English may not judge B's utterances in (\ref{ai}a) and (\ref{ai}b) to be acceptable and prefer versions with {\em enough}: {\em She was smart enough to buy one the day they went on sale}. We hypothesize that such judgments of unacceptability are due to such speakers dispreferring EAS in which the prejacent is at-issue. Given the hypothesis we advance below, that the prejacent projects to the extent that it is not at-issue, we expect such speakers to also judge NEAS like (\ref{nat}) in which the prejacent does not project to be unacceptable. Crucially, however, even speakers who prefer to realize (\ref{ai}a) and (\ref{ai}b) with {\em enough} can retrieve the intended interpretations of the variants without {\em enough}.} Thus, in (\ref{ai}a) and (\ref{ai}b), the prejacent is at-issue and the generalization, that in buying a ticket the day the tickets went on sale Sam is be smart, is not at-issue. The Discourse Questions in (\ref{ai}c) and (\ref{ai}d), on the other hand, are not about the prejacent:  both the question of how Sam got a ticket in (\ref{ai}c) and whether Sam was smart to buy a ticket the day they went on sale in (\ref{ai}d) are not answered by the prejacent of B's utterances, that Sam got a ticket, but by the generalization, that in buying a ticket the day the tickets went on sale Sam is smart. Thus, in (\ref{ai}c) and (\ref{ai}d), the generalization is at-issue and the prejacent is not at-issue.

@@

Identifying which utterance content is at-issue is critical not only to understanding what the utterance is about, but also because of the connection between at-issueness and projectivity (e.g., \citealt{potts05,brst-salt10,best-question,brst-ar}). For EAS, identifying whether the prejacent or the generalization are at-issue thus also plays a role in identifying whether the prejacent projects, as we discuss next.

@@

To illustrate the proposal, consider the examples in (\ref{ai2}a) and (\ref{ai2}b), where B utters NEASs. The prejacent is at-issue in (\ref{ai2}a) because A's interrogative utterance, assumed to be the Discourse Question, is about the prejacent. Thus, according to our proposal, the prejacent of B's utterance in (\ref{ai2}a) is predicted to not project: accordingly, B's utterance means that Sam didn't buy a ticket the day they went on sale (the prejacent does not project) and that buying a ticket the day they went on sale would have been smart for Sam. In (\ref{ai2}b), on the other hand, A's interrogative utterance, is about the generalization. Thus, the prejacent of B's utterance is not at-issue and, according to our proposal, is predicted to project. Accordingly, B's utterance is interpreted to mean that Sam bought a ticket the day they went on sale and that Sam buying a ticket the day they went on sale wasn't smart.

@@

\begin{xlist}
\ex
\begin{xlist}
\exi{A:} The show was sold out. Did Sam get a ticket?
\exi{B:} She wasn't smart to buy one the day they went on sale. (So, she didn't go to the show.)
\end{xlist}

\exi{B:} Sam wasn't smart to buy one the day they went on sale. The price indeed went down after a few days.

@@

As mentioned above, we take projectivity to be a gradient property of utterance content rather than a binary, categorical one. This assumption is motivated by experimental findings reported on in \citealt{tbd-variability} as well as the corpus study reported on in section \ref{s2}, which showed that authors can be taken to be committed to the prejacent of NEASs to varying degrees. Thus, we assume that what it means for projectivity to be a gradient property is that a reader's (or listener's) judgment that a content is projective to a certain extent means that the reader takes the author (or speaker) to be committed to the content to that extent. On this interpretation, projectivity being a gradient property is a consequence of author (or speaker) commitment being a gradient property (for discussion, see \citealt{tbd-variability}). Accordingly, to capture the relation between projectivity and not-at-issueness, we assume \citepos{tbd-variability} Gradient Projection Principle:\footnote{The question of why not-at-issue content projects has received several different answers. \citet{brst-salt10} proposed that not-at-issue content is not targeted by operators, such as negation, thereby resulting in the content projecting over such operators. In this paper, we follow \citet{abrusan2011} in assuming that the not-at-issue content is backgrounded and projects as a result of its discourse status.}

@@

Preliminary evidence that the projectivity of the prejacent of EASs depends on its at-issueness comes from \citealt{tbd-variability}, who investigated the Gradient Projection Principle on the basis of 19 projective contents associated with American English expressions. \citet{tbd-variability} collected both projectivity and at-issueness ratings for these 19 contents, one of which was the prejacent of EASs with {\em stupid} embedded under the polar question operator. The four evaluative adjective polar questions studied in their work are given in (\ref{stupid2}):

@@

\ex Is Richie stupid to be a stuntman? \hfill (\citealt{tbd-variability}, Appendix A)

@@

\caption{Projectivity ratings against not-at-issueness ratings for the prejacent of the four polar question EASs in \citealt{tbd-variability}. Each dot represents one participant's ratings. Linear smoothers with 95\% confidence intervals overlaid.}
\label{f-corr}
\end{figure}

In \citepos{tbd-variability} experiment, the evaluative adjective polar questions in (\ref{stupid2}) were not presented as responses to interrogative utterances that made the Discourse Questions explicit. Rather, participants were asked to imagine that the polar questions were uttered by speakers who they overhear at a party. The absence of an explicit Discourse Question is, of course, not unusual and rather the norm in naturally occurring data. Given the importance of the Discourse Question to the interpretation of EASs, in particular the at-issueness of the prejacent, we need to understand how the Discourse Question of utterances not presented in response to an interrogative utterance is constrained. In this paper, we consider two cues to the Discourse Questions of utterances of EASs: the information structure of the utterance (section \ref{s331}) and the discourse status of the generalization (section \ref{s332}).\footnote{\citet{abrusan2011} proposed that entailments of utterances that are about the running time of the main event are the default main point, i.e., what we have referred to as the at-issue content. However, given Abrus\'an's notion of aboutness, both the prejacent and the generalization of EASs are about the running time of the main event. For instance, \citet[508]{abrusan2011} took the entailment of (i), that John solved the exercise, to be ``non-accidentally (i.e. necessarily) about the matrix event time''. By the same argument, the prejacent of (ii) would be about the matrix event time, which means that neither the prejacent nor the generalization are predicted by \citealt{abrusan2011} to be the default at-issue content of EASs.

@@

Question-answer congruence is straightforwardly modeled in an
alternative semantics framework (see, e.g., \citealt{rooth85,rooth92}), which
analyzes both the focus semantic meaning of the answer and the meaning of
the question in terms of alternatives. In such a framework, the meanings
of the questions in (\ref{qa1}) and (\ref{qa2}) are sets of propositions
that are answers to the questions, as shown in (\ref{cong1}a) and
(\ref{cong2}a), respectively. The focus
semantic meanings of the answers in (\ref{qa1}J) and (\ref{qa2}J) are
focus alternatives sets, i.e., sets of propositions derived by
abstracting over the focused expressions, as shown in (\ref{cong1}b) and
(\ref{cong2}b).

@@

\ex \sem{(\ref{qa1}M)}{M,g} = \{David likes Turkish coffee, David likes Lebanese coffee, David likes Irish coffee\}

\ex \sem{(\ref{qa1}J)}{M,g} = \{David likes Turkish coffee, David likes Lebanese coffee, David likes Irish coffee, David likes cold coffee,\ldots\}

@@

\ex \sem{(\ref{qa2}M)}{M,g} = \{David likes Turkish coffee, Craige likes Turkish coffee, Mandy likes Turkish coffee, Judith likes Turkish coffee\}

\ex \sem{(\ref{qa2}J)}{M,g} = \{David likes Turkish coffee, Craige likes Turkish coffee, Mandy likes Turkish coffee, Judith likes Turkish coffee, Adam likes Turkish coffee,\ldots\}

@@

\ex CQ of (\ref{np}a), a subset of (\ref{np}b), entails that somebody bought a ticket the day they went on sale and that, in buying a ticket the day they went on sale, somebody was smart.

@@

To illustrate how the Current Question constrains the Discourse Question, consider an utterance of the example in (\ref{aib}), which is presented with its focus alternative set. The Current Question, as a subset of this focus alternative set, entails that somebody bought a ticket the day they went on sale and that somebody buying a ticket the day they went on sale is smart. That is, when this utterance is made, what is already taken to be part of the common ground is that somebody bought a ticket the day they went on sale and that, in buying a ticket the day they went on sale, somebody was smart. Only some Discourse Questions are plausible, given this common ground information. In particular, both the Discourse Questions `Who got a ticket?' and `Did Sam get a ticket?' are plausible because both are unanswered given the entailments of the Current Question. On the other hand, Discourse Questions like `How did Sam get a ticket?' and `Was Sam smart to buy a ticket the day they went on sale?' are less plausible because what is common ground already answers these Discourse Questions.

@@

In sum, this section has illustrated how the focus marking of an utterance of an EAS constrains the Discourse Question of the utterance and, thereby, the at-issueness of the prejacent. For EASs that are embedded under an entailment-canceling operator, the focus marking of utterances of such EAS is thereby predicted to influence the projectivity of the prejacent.

@@

\ex\label{principle} {\bf Gradient non-redundancy principle for at-issue content} \\ For utterance content $c$, the stronger the inference to the truth of $c$ from the common ground, the less likely is it that the Discourse Question of the utterance is about $c$, i.e., $c$ is less at-issue.

@@

\end{xlist}
\end{exe} 
If the evaluative adjectives of these NEASs are focused, the high mean projectivity ratings of the prejacents are expected, given that NEAS with focused evaluative adjectives address Discourse Questions according to which the prejacent is not at-issue.

>>>>

EASs have traditionally been analyzed as presupposing the prejacent (e.g., \citealt{norrick78,barker02,oshima09b,kertz2010}). On such analyses,  evaluative adjectives like {\em stupid} lexically specify that the prejacent
must be entailed by or satisfied in the common ground of the interlocutors in
order for an EAS to be interpretable (e.g.,
\citealt{heim83,vds92}). Such analyses are motivated by and straightforwardly account for examples like those in (\ref{f2}a-d) in which the prejacent projects over entailment-canceling operators, i.e., negation in (\ref{f2}a), a polar question in (\ref{f2}b), the possibility adverb {\em perhaps} in (\ref{f2}c) and the antecedent of the conditional in (\ref{f2}d). That is, speakers who utter (\ref{f2}a-d) may be taken to be committed to the prejacent, that Feynman danced on the table, even though the evaluative adjective occurs under an entailment-canceling operator. The evaluation, on the other hand, that Feynman dancing on the table was stupid, is not projective content but denied in (\ref{f2}a), questioned in (\ref{f2}b), and merely considered a possibility in (\ref{f2}c) and (\ref{f2}d).

@@

We discuss below why such examples are problematic for analyses according to which the prejacent is a presupposition. But we first need to mention that there is variation in the population of native speakers of American English: whereas all such speakers are able to interpret EASs in which the prejacent does not project, i.e., are able to retrieve the intended interpretations of examples like (\ref{nat}), a sizeable portion strongly prefers to realize such interpretations with variants that includes {\em enough}, as in {\em I wasn't stupid enough to go stumbling through the junkyard in the dark} for (\ref{nat}b); see also \citealt{karttunen2013} and \citealt{karttunen-etal2014} for this observation. The ability of native speakers of American English to retrieve not just the projecting but also the non-projecting interpretation is necessitated by  the fact that EASs in which the prejacent does not project are part of American English, as evidenced by the existence of naturally occurring examples like (\ref{nat}); further  such examples are presented in this paper. Additional evidence comes from the fact that there are native speakers who judge EASs in which the prejacent does not project to be perfectly acceptable\footnote{The audiences to which this research was presented over the years always included native speakers that judged examples like (\ref{nat}) to be acceptable. More systematic evidence for the existence of such speakers comes from the acceptability rating study presented in Appendix \ref{s-acc}: of the 94 self-reported native speakers of American English that participated in the study, about 20-30\% judged negated evaluative adjective sentences without {\em enough} to be acceptable under a non-projecting interpretation of the prejacent.} and who produce examples like (\ref{nat}). In short, given that EASs in which the prejacent does not project are part of American English, we assume that all native speakers of American English are able to retrieve both interpretations even if they might not produce EASs in which the prejacent does not project. Our goal in this paper is to analyze the interpretation of EASs; we return to the observed variation among native speakers of American English in section \ref{s5}.

@@

This behavior of the evaluation is not predicted by the two analyses that consider content other than the prejacent: \citealt{oshima09b} and \citealt{barker02}. In \citealt{oshima09b}, the prejacent is a presupposition and what is asserted is the following complex content: from the prejacent it can be inferred that the denotation of the subject is in the extension of the evaluative adjective (p.371). That is, (\ref{f2}a) is correctly predicted to presuppose that Feynman danced on the table and to assert that it cannot be inferred from Feynman dancing on the table that Feynman is stupid. Similarly then, (\ref{nat}b) is predicted to convey that the speaker did not stumble through the junkyard (the prejacent is locally accommodated) and to assert that it cannot be inferred from the speaker stumbling through the junkyard that the speaker is stupid. This assertion is too weak: crucially, it does not follow the speaker stumbling through the junkyard would be stupid.

@@

\citepos{karttunen-etal2014} analysis improves over previous analyses because it is able to derive both attested interpretations. But the analysis comes at a cost: first, evaluative adjectives are systematically ambiguous, and, second, as we argue in section \ref{s3}, it misses out on an empirical generalization about how common ground information  influences the interpretation of EASs. {\bf  IS IT REALLY THE CASE THAT THEY MISS OUT ON THIS GENERALIZATION: IF CG SPECIFIES THAT IT IS FORTUNATE TO WIN THE LOTTERY, THEN ONE LEXICAL ENTRY IS MORE LIKELY} In this paper, we develop an analysis of EASs that accounts for the two interpretations with a single lexical entry for evaluative adjectives. The crux of the analysis is that evaluative adjectives do not lexically specify that the prejacent is a presupposition. In section \ref{s2}, we show, based on a study of naturally occurring NEASs, that the prejacent is not highly projective. We take this to provide empirical support for not analyzing the prejacent as a lexically-specified presupposition. In section \ref{s3}, we present an analysis of EASs according to which the projectivity of the prejacent of an utterance in which the evaluative adjective is embedded under an entailment-canceling operator is predicted by the extent to which the prejacent is at-issue with respect to the Discourse Question addressed by the utterance. We test three predictions of the analysis in section \ref{s4} and discuss two open issues in section \ref{s5}. The paper concludes in section \ref{s6}.

{\bf We now clarify in the introduction that we develop a question-based analysis of EAS to capture projection variability observed for the prejacent of EAS, but that it is motivated to conventionally specify the projectivity of other content.}

@@

{\bf The main point here is to show that the prejacent is not very projective, at least in the corpus study we did. This motivates an analysis as not lexically specified presupposition. Only present graph and examples in section 1, move rest to appendix.}

\citet{karttunen-etal2014} provided naturally occurring examples that showed that the prejacent of NEASs need not project. The goal of the study we present in this section was to more systematically explore how projective the prejacent is in naturally occurring NEAS. Following AUTHORS 2018, we assume that projectivity is a gradient property of utterance content: the more the speaker or author of the utterance is taken to be committed to the truth of an utterance content, the more projective the content is. {\bf Elaborate here, along the lines of the JoS paper, why projectivity is assumed to be gradient and what that means. Elaborate here or later that at-issueness is also a gradient property of utterance content.} To explore the projectivity of the prejacent, native speakers of American English were presented with naturally occurring NEAS and asked to rate the projectivity of the prejacent.\footnote{\label{f-git}The
materials, data and the R code for generating the figures and analyses of the experiments reported on in this paper are available at {\tt [name of GitHub repository]}.}
%\url{https://github.com/judith-tonhauser/evaluative-adjectives}.}

@@

Each target stimulus was attributed to an author who was identified by name, as shown for the sample stimuli in (\ref{target1}). As shown in the `Question to participants' presented with each stimulus in (\ref{target1}), we used the `certain that' diagnostic for projectivity to assess whether participants took the author to be committed to the prejacent, i.e., whether the prejacent projects over negation. For other applications of the `certain that' diagnostic for projection see \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017} and AUTHORS 2018. In these `Questions to participants', the prejacent was realized as the finite complement clause of {\em certain}. The tense of the finite clause that realized the prejacent was determined by the temporal and aspectual properties of the
NEAS: it was past tense for past tense NEAS, as in (\ref{target1}a), and future tense for future tense NEAS, as in (\ref{target1}b); for non-past tense NEAS, the finite clause that realized the prejacent was realized in the non-past tense if the eventuality denoted by the prejacent was
stative, as in (\ref{target1}c), and a disjunction of a past tense and a future tense verb if the prejacent was eventive, as in (\ref{target1}d).

@@

In sum, our study of naturally occurring NEASs shows that the prejacent is not highly projective. These findings converge with those of AUTHORS (2018), who studied the projectivity of a variety of utterance contents and found that the prejacent of four (constructed) polar interrogative EASs with {\em stupid} was significantly less projective than the content of the clausal complement of the emotive predicate {\em be annoyed} and the cognitive predicate {\em know}. In the next section, we develop an analysis according to which the projectivity of the prejacent is not due to its lexical status as a presupposition but instead derived from the question that the utterance of the EAS is taken to address.

@@

{\bf say that both prejacent and generalization can be new information, i.e., not associated with SCF}

{\bf clarify in this section that these are entailments but they may still project, i.e., they are not regular entailments. also point to assumption in literature that some presuppositions are entailed by affirmative sentences. And Abrusan's analysis and Anand and Hacquard's assumption}

@@

\ex {\bf give more precise characterization of prejacent and generalization, e.g., using logical formulas and drawing on Barker's semantics for the evaluative adjectives}

Feynman was stupid to dance on the table

Generalization: $stupid'(dance.on.table', f)$ ``the degree to which Feynman dancing on the table at some time is stupid is higher than the contextual standard for {\em stupid}''

Prejacent: $dance.on.table'(f,rt)$

Evaluation: $dance.on.table'(f,rt) \wedge stupid'(dance.on.table', f)$, i.e., Feynman was stupid to dance on the table

@@

The examples in (\ref{ai}) illustrate four Discourse Questions that utterances of EASs can address. These Discourse Questions show that both lexical entailments of EASs, the prejacent and the generalization, can be at-issue. The Discourse Questions in (\ref{ai}a) and (\ref{ai}b) are about the prejacent: both the question of whether Sam got a ticket in (\ref{ai}a) and who got a ticket in (\ref{ai}b) are answered by the prejacent of B's utterances, that Sam got a ticket.\footnote{Some native speakers of American English would prefer to produce B's utterances in (\ref{ai}a) and (\ref{ai}b) with {\em enough}: {\em She was smart enough to buy one the day they went on sale}. We hypothesize that such judgments of unacceptability are due to such speakers dispreferring EAS in which the prejacent is at-issue. Given the hypothesis we advance below, that the prejacent projects to the extent that it is not at-issue, we expect such speakers to also judge NEAS like (\ref{nat}) in which the prejacent does not project to be unacceptable. Crucially, as discussed in section \ref{s1}, even speakers who prefer to realize (\ref{ai}a) and (\ref{ai}b) with {\em enough} can retrieve the intended interpretations of the variants without {\em enough}.} Thus, in (\ref{ai}a) and (\ref{ai}b), the prejacent is at-issue and the generalization, that in buying a ticket the day the tickets went on sale Sam is be smart, is not at-issue. The Discourse Questions in (\ref{ai}c) and (\ref{ai}d), on the other hand, are not about the prejacent:  both the question of how Sam got a ticket in (\ref{ai}c) and whether Sam was smart to buy a ticket the day they went on sale in (\ref{ai}d) are not answered by the prejacent of B's utterances, that Sam got a ticket, but by the generalization, that in buying a ticket the day the tickets went on sale Sam is smart. Thus, in (\ref{ai}c) and (\ref{ai}d), the generalization is at-issue and the prejacent is not at-issue.

@@

Identifying which utterance content is at-issue is critical not only to understanding what the utterance is about, but also because of the connection between at-issueness and projectivity (e.g., \citealt{potts05,brst-salt10,best-question,brst-ar}, AUTHORS 2018). For EAS, identifying whether the prejacent or the generalization are at-issue thus also plays a role in identifying whether the prejacent projects, as we discuss next.

@@

To illustrate the proposal, consider the minimal pair in (\ref{ai2}a) and (\ref{ai2}b), where B utters the same NEAS, modulo whether the subject is a pronoun or a proper name. The prejacent of this NEAS is at-issue in (\ref{ai2}a) because A's interrogative utterance, assumed to be the Discourse Question, is about the prejacent. Thus, according to our proposal, the prejacent of B's utterance in (\ref{ai2}a) is predicted to not project: accordingly, B's utterance means that Sam didn't buy a ticket the day they went on sale (the prejacent does not project) and that buying a ticket the day they went on sale would have been smart for Sam. In (\ref{ai2}b), on the other hand, A's interrogative utterance, is about the generalization. Thus, the prejacent of B's utterance is not at-issue and, according to our proposal, is predicted to project. Accordingly, B's utterance is interpreted to mean that Sam bought a ticket the day they went on sale and that Sam buying a ticket the day they went on sale wasn't smart.

@@

\begin{xlist}
\ex
\begin{xlist}
\exi{A:} The show was sold out. Did Sam get a ticket?
\exi{B:} She wasn't smart to buy a ticket the day they went on sale. (So, she didn't go to the show.)
\end{xlist}

\ex
\begin{xlist}

\exi{A:} You keep criticizing Sam for doing some not-so-smart things. I'm not sure I agree. Can you give me an example (of something not-so-smart that Sam did)?

\exi{B:} Sure! Remember the rock show we talked about? Sam wasn't smart to buy a ticket the day they went on sale. The price went down a couple of days later.

%\exi{A:} Was Sam smart to buy a ticket for the show the day the tickets went on sale? The price would have gone down after a few days!
%
%\exi{B:} Sam wasn't smart to buy one the day they went on sale. The price indeed went down after a few days.

@@

As mentioned above, we take projectivity to be a gradient property of utterance content rather than a binary, categorical one. This assumption is motivated by experimental findings reported on in AUTHORS 2018 as well as the corpus study reported on in section \ref{s2}, which showed that authors can be taken to be committed to the prejacent of NEASs to varying degrees. Thus, we assume that what it means for projectivity to be a gradient property is that a reader's (or listener's) judgment that a content is projective to a certain extent means that the reader takes the author (or speaker) to be committed to the content to that extent. On this interpretation, projectivity being a gradient property is a consequence of author (or speaker) commitment being a gradient property (for discussion, see AUTHORS 2018). Accordingly, to capture the relation between projectivity and not-at-issueness, we assume AUTHORS' 2018 Gradient Projection Principle: {\bf SAY MORE HERE ABOUT HOW THIS PRINCIPLE IS MOTIVATED, TO ADDRESS REVIEWER}\footnote{{\bf move this into the main body and also mention potts 2005.} The question of why not-at-issue content projects has received several different answers. \citet{brst-salt10} proposed that not-at-issue content is not targeted by operators, such as negation, thereby resulting in the content projecting over such operators. In this paper, we follow \citet{abrusan2011} in assuming that the not-at-issue content is backgrounded and projects as a result of its discourse status.}

@@

Preliminary evidence that the projectivity of the prejacent of EASs depends on its at-issueness comes from AUTHORS 2018, who investigated the Gradient Projection Principle on the basis of 19 projective contents associated with American English expressions. AUTHORS (2018) collected both projectivity and at-issueness ratings for these 19 contents, one of which was the prejacent of EASs with {\em stupid} embedded under the polar question operator. The four evaluative adjective polar questions studied in their work are given in (\ref{stupid2}):

@@

\ex Is Richie stupid to be a stuntman? \hfill (AUTHORS 2018, Appendix A)

@@

\caption{Projectivity ratings against not-at-issueness ratings for the prejacent of the four polar question EASs in AUTHORS 2018. Each dot represents one participant's ratings. Linear smoothers with 95\% confidence intervals overlaid.}
\label{f-corr}
\end{figure}

In AUTHORS' 2018 experiment, the evaluative adjective polar questions in (\ref{stupid2}) were not presented as responses to interrogative utterances that made the Discourse Questions explicit. Rather, participants were asked to imagine that the polar questions were uttered by speakers who they overhear at a party. The absence of an explicit Discourse Question is, of course, not unusual but rather the norm in naturally occurring data. Given the importance of the Discourse Question to the interpretation of EASs, in particular the at-issueness of the prejacent, we need to understand how the Discourse Question of utterances not presented in response to an interrogative utterance is constrained. In this paper, we consider two cues to the Discourse Questions of utterances of EASs: the information structure of the utterance (section \ref{s331}) and the discourse status of the generalization (section \ref{s332}).\footnote{\citet{abrusan2011} proposed that entailments of utterances that are about the running time of the main event are the default main point, i.e., what we have referred to as the at-issue content. However, given Abrus\'an's notion of aboutness, both the prejacent and the generalization of EASs are about the running time of the main event. For instance, \citet[508]{abrusan2011} took the entailment of (i), that John solved the exercise, to be ``non-accidentally (i.e. necessarily) about the matrix event time''. By the same argument, the prejacent of (ii) would be about the matrix event time, which means that neither the prejacent nor the generalization are predicted by \citealt{abrusan2011} to be the default at-issue content of EASs.

@@

Question-answer congruence is straightforwardly modeled in an
alternative semantics framework (see, e.g., \citealt{rooth85,rooth92}), which
analyzes both the focus semantic meaning of the answer and the meaning of
the question in terms of alternatives. In such a framework, the (ordinary semantic) meanings
of the questions in (\ref{qa1}) and (\ref{qa2}) are sets of propositions
that are answers to the questions, as shown in (\ref{cong1}a) and
(\ref{cong2}a), respectively (where the superscript {\em o} indicates `ordinary semantic meaning'). The focus
semantic meanings of the answers in (\ref{qa1}J) and (\ref{qa2}J) are
focus alternatives sets, i.e., sets of propositions derived by
abstracting over the focused expressions, as shown in (\ref{cong1}b) and
(\ref{cong2}b), respectively (where the superscript {\em f} indicates `focus semantic meaning').

@@

\ex \sem{(\ref{qa1}M)}{o} = \{David likes Turkish coffee, David likes Lebanese coffee, David likes Irish coffee\}

\ex \sem{(\ref{qa1}J)}{f} = \{David likes Turkish coffee, David likes Lebanese coffee, David likes Irish coffee, David likes cold coffee,\ldots\}

@@

\ex \sem{(\ref{qa2}M)}{o} = \{David likes Turkish coffee, Craige likes Turkish coffee, Mandy likes Turkish coffee, Judith likes Turkish coffee\}

\ex \sem{(\ref{qa2}J)}{f} = \{David likes Turkish coffee, Craige likes Turkish coffee, Mandy likes Turkish coffee, Judith likes Turkish coffee, Adam likes Turkish coffee,\ldots\}

@@

\ex CQ of (\ref{np}a), a subset of (\ref{np}b), entails that somebody bought a ticket the day they went on sale and that, in buying a ticket the day they went on sale, somebody was smart. {\bf THAT PERSON WAS SMART?}

@@

Discourse questions
\begin{itemize}
\exi{A:} The show was sold out. Did Sam get a ticket?

\exi{A:} There were so few tickets for the show! Who got a ticket?

\exi{A:} The tickets for the show sold out really quickly! How did Sam get a ticket?

\exi{A:} Was Sam smart to buy a ticket for the show the day the tickets went on sale? The price would have gone down after a few days!
\end{itemize}

To illustrate how the Current Question constrains the Discourse Question, consider an utterance of the example in (\ref{aib}), which is presented with its focus alternative set. The Current Question, as a subset of this focus alternative set, entails that somebody bought a ticket the day they went on sale and that somebody buying a ticket the day they went on sale is smart. {\bf THAT PERSON WAS SMART?}  That is, when this utterance is made, what is already taken to be part of the common ground is that somebody bought a ticket the day they went on sale and that, in buying a ticket the day they went on sale, somebody was smart. {\bf THAT PERSON WAS SMART?} Only some Discourse Questions are plausible, given this common ground information. In particular, both the Discourse Questions `Who got a ticket?' and `Did Sam get a ticket?' are plausible because both are unanswered given the entailments of the Current Question. On the other hand, Discourse Questions like `How did Sam get a ticket?' and `Was Sam smart to buy a ticket the day they went on sale?' are less plausible because what is common ground already answers these Discourse Questions.

@@

In sum, this section  illustrated how the focus marking of an utterance of an EAS constrains the Discourse Question of the utterance and, thereby, the at-issueness of the prejacent. For EASs that are embedded under an entailment-canceling operator, the focus marking of utterances of such EAS is thereby predicted to influence the projectivity of the prejacent.

@@

\ex\label{principle} {\bf Gradient non-redundancy principle for at-issue content} \\ For utterance content $c$, the stronger the inference to the truth of $c$ from the common ground, the less likely is it that the Discourse Question of the utterance is about $c$, i.e., $c$ is less at-issue. {\bf ADD NOTE TO ENSURE THAT THIS, TOO, IS ONLY UNDERSTOOD AS A CONDITIONAL, NOT BI-CONDITIONAL}

@@

\end{xlist}
\end{exe} 
If the evaluative adjectives of these NEASs are focused, the high mean certainty ratings of the prejacents are expected, given that NEAS with focused evaluative adjectives address Discourse Questions according to which the prejacent is not at-issue.
